### 2018-01-05 includes/analysis.tex
Added paragraph stubs.

\section{Experiments}

    Here we present instances we generated, and our empirical results.

    \subsection{Concats-Extracts}

        Here we describe the problem. CVC4 did poorly. Z3str3 did well.

    \subsection{Regex Suite}

        Here we describe the problem. CVC4 did poorly. Z3str3 did well.

    \subsection{Lazy-Hard}

        Here we describe the problem. CVC4 did poorly. Z3str3 did well.

    \subsection{Eager-Hard}

        Here we describe the problem. CVC4 did well. Z3str3 did poorly.

\section{Analysis}

    Here we talk about Lazy-Hard and Eager-Hard, and how the algorithms of CVC4 and Z3str3 differ, eliciting the observed behaviour.

    \subsection{Lazy-Hard}

        Here is why CVC4 does poorly and Z3str3 does well.

    \subsection{Eager-Hard}

        Here is why CVC4 does well and Z3str3 does poorly.