### 2018-01-14 includes/recommendation.tex
Applying first round of feedback.

\section{\us{} Recommendation}

    This section outlines a change for \us{}, which we believe will improve its performance on the Eager-Hard problem set and similar problems. We recommend a partial order treatment of equivalence class comparisons. First, we describe a partial order of equivalence classes. Then, we use it in equivalence class comparisons and more closely approximate equality between them.

    % TODO:
    % - definition of partial order
    % - definition of current class relation
    % - definition of class relation under partial order