### 2018-01-20 includes/prior.tex
Addressing Federico's feedback.

\section{Related Work}

    Naturally, many solver developers author their own unit test suites to validate their solvers \cite{cvc4-tests} \cite{z3str3-tests} \cite{z3str2-tests}. In addition, several popular problem suites are publicly available for solver validation, such as the Kaluza \cite{kaluza} and Kausler \cite{kausler} suites.

    There are likewise several fuzzers and problem generators currently available, but none of them support string or regex generation or transformation. For example the FuzzSMT\cite{fuzzsmt} tool generates \smt{} problems with bit vectors and arrays, but does not support strings or regexes. The SMTpp\cite{smtpp} tool preprocesses and simplifies problems, but does not generate new ones or transform existing ones.

% FuzzSMT: http://fmv.jku.at/papers/BrummayerBiere-SMT09.pdf
% SMTpp: http://www.verit-solver.org/papers/smt2015.pdf