### 2018-01-27 includes/problems.tex
Adding more sections. Adding references.

\section{Problem Suites}
\label{sec:problems}

    We generated several suites of string and regex problems using \fuzzer{}. \footnote{For review anonymity, we removed the link to the problem repository. It will be provided separately in the artifact submission track and will appear in the final version of the paper.} Table~\ref{tbl:generated} describes the \texttt{generated} suite that we generated using \generator{}.

    \begin{table}
        \centering
        \begin{tabular}{|l|l|}
            \hline
            \textbf{Name}             & \textbf{Problems have ...} \\ \hline
            \textit{Concats}          & A single, deep tree of concats. \\ \hline
            \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
            \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
            \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
            \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
            \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
            \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
            \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
            \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
            \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
        \end{tabular}
        \caption{\generator{}-generated problem sets.}
        \label{tbl:generated}
    \end{table}

    The \texttt{transformed} suite contains benchmarks derived from existing seed problems by iteratively applying \transformer{}. Every problem in the suite is named according to its seed and the transformations it undertook. For example, a hypothetical file named ``seed-fuzz-graft.smt2'' began as a file named ``seed.smt'' and was transformed twice: first by fuzz, and then by graft. The suite is divided into two categories. For an overview see Table~\ref{tbl:transformed}.

    \begin{table}
        \centering
        \begin{tabular}{|l|l|r|}
            \hline
            \textbf{Name}   & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
            \textit{Amazon} & Two industrial regex membership problems & 516\\ \hline
            \textit{Regex}  & Z3str2 regular expression test suite     & 10,577\\ \hline
        \end{tabular}
        \caption{\transformer{}-generated problem sets.}
        \label{tbl:transformed}
    \end{table}

    The first category, labeled Amazon, started as two seed files.
    The first seed, z3-regex-1.smt2, encodes the assertions ``S not in aaabbb*'' and ``S in aaa*bbb*'' for a string constant S,
    while the second seed, z3-regex-2-negative.smt2, encodes the assertions ``S in aaabbb*'' and ``S not in aaa*bbb*''.
    Both problems come from our industrial collaborators.
    In total, 516 unique problems were generated from these two seeds.

    The second category, Regex, is seeded by the z3str2 regular expression test suite \cite{z3str2-tests}.
    This test suite contains 42 problems dealing with a wide variety of regular expressions.
    Through cumulative transformations, as before, we were able to produce 10,577 unique benchmarks.

    \subsection{Usefulness to \us{}}

        This section describes the bugs we discovered in \us{} (that were subsequently fixed) by testing it with the \fuzzer{} problem suites.

        The first bug we found caused \us{} to perform poorly when equivalence classes of arithmetic terms was very large (e.g. where many strings were equal to the empty string, and thus had length 0). We discovered this bug using the \textit{Concats} problem suite. The slowdown was caused by a loop over the equivalence class of these terms to search for an integer constant. The fix was to check the root of the equivalence class for the integer constant, as Z3 makes "interpreted terms" (i.e. constants) the root of equivalence classes whenever possible. This bug was last observed in commit 6308636, and was fixed in commit 3865c45.

        The second bug we found caused \us{} to perform poorly when a variable had a large length constraint. We discovered this bug using the \textit{Lengths} problem suite. The slowdown was caused by the solver performing a linear search for a satisfying length when producing a model for the variable. This bug was an inconsistency between the implementation and the behaviour described in the \us{} paper \cite{z3str3}. This bug was last observed in commit 66bc68f, and was fixed in commit 7b536e9.

        \todo{Federico - Can we describe the floating point error we found?}

        \todo{Dmitry: maybe show graphs with before and after?}