### 2018-01-29 includes/problems.tex
a pass

\section{Problem Suites}
\label{sec:problems}
\vspace{-0.1in}

\subsection{Generated suite}
\vspace{-0.1in}
    We generated several suites of string and regex problems using \fuzzer{}\footnote{For review anonymity, we removed the link to the source code. It will be provided separately in the artifact submission.}. Table~\ref{tbl:generated} describes the suite \texttt{generated} by \generator{}.

\begin{table}[h]
\vspace{-0.4in}
\centering
\caption{\generator{}-generated problem sets.}
\label{tbl:generated}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.8\columnwidth}{!}{
        \begin{tabular}{|l|l|}
            \hline
            \textbf{Name}             & \textbf{Problems have ...} \\ \hline
            \textit{Concats}          & A single, deep tree of concats. \\ \hline
            \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
            \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
            \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
            \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
            \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
            \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
            \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
            \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
            \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
        \end{tabular}  
  }
  \egroup
  \vspace{-0.2in}
}
\end{table}

%     \begin{table}
%         \centering
%         \begin{tabular}{|l|l|}
%             \hline
%             \textbf{Name}             & \textbf{Problems have ...} \\ \hline
%             \textit{Concats}          & A single, deep tree of concats. \\ \hline
%             \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
%             \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
%             \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
%             \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
%             \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
%             \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
%             \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
%             \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
%             \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
%         \end{tabular}
%         \caption{\generator{}-generated problem sets.}
%         \label{tbl:generated}
%     \end{table}

\vspace{-0.2in}
\subsection{Transformed suite}
\vspace{-0.1in}
    The \texttt{transformed} suite contains benchmarks derived from existing seed problems by iteratively applying \transformer{}. Table~\ref{tbl:transformed} shows the overview. Every problem in the suite is named according to its seed and the transformations it undertook. For example, input \texttt{seed-fuzz-graft.smt2} was transformed by applying \texttt{fuzz} and then \texttt{graft} from \texttt{seed.smt2}.

\begin{table}[h]
\vspace{-0.35in}
\centering
\caption{Transformed problem sets generated by \transformer{}.}
\label{tbl:transformed}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.6\columnwidth}{!}{
          \begin{tabular}{|l|l|r|}
            \hline
            \textbf{Name}   & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
            \textit{Amazon} & Two industrial regex membership problems & 516\\ \hline
            \textit{Regex}  & Z3str2 regular expression test suite     & 10,577\\ \hline
        \end{tabular}
    }
  \egroup
  \vspace{-0.2in}
}
\end{table}

%     \begin{table}
%         \centering
%         \begin{tabular}{|l|l|r|}
%             \hline
%             \textbf{Name}   & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
%             \textit{Amazon} & Two industrial regex membership problems & 516\\ \hline
%             \textit{Regex}  & Z3str2 regular expression test suite     & 10,577\\ \hline
%         \end{tabular}
%         \caption{\transformer{}-generated problem sets.}
%         \label{tbl:transformed}
%     \end{table}

    The Amazon category is obtained from two seeds: (1) \texttt{z3-regex-1.smt2} encodes the assertions \texttt{(S not in aaabbb*)} and \texttt{(S in aaa*bbb*)} for a string constant S; (2) \texttt{z3-regex-2-negative.smt2} includes regex \texttt{(S in aaabbb*)} and \texttt{(S not in aaa*bbb*)}.
    Both problems come from our industrial collaborators. In total, 516 unique problems were generated from these two seeds.

    The Regex category is seeded by the z3str2 regular expression test suite \cite{z3str2-tests}.
    This test suite contains 42 problems dealing with a wide variety of regular expressions.
    Through cumulative transformations, we have $10,577$ unique benchmarks.