### 2018-01-29 includes/prior.tex
one pass all the way to the end. Conclusion and related work need work.

\section{Related Work}

    Naturally, many solver developers author their own test suites to validate their solvers 
    \cite{cvc4-tests} \cite{z3str3-tests} \cite{z3str2-tests}. 
    In addition, several popular problem suites are publicly available for solver validation, 
    such as the Kaluza \cite{kaluza} and Kausler \cite{kausler} suites.

    There are likewise several fuzzers and problem generators currently available.
    However, none of these fuzzers can generate and transform string and regex problems. 
    \todo{I know we talked about this one before, but I'm confused by it again.}
    For example the FuzzSMT\cite{fuzzsmt} tool generates \smt{} problems with bit vectors and arrays, 
    but does not support strings or regexes. 
    The SMTpp\cite{smtpp} tool pre-processes and simplifies problems, but does not generate new ones or fuzz existing ones.

% FuzzSMT: http://fmv.jku.at/papers/BrummayerBiere-SMT09.pdf
% SMTpp: http://www.verit-solver.org/papers/smt2015.pdf