### 2018-01-30 includes/prior.tex
editing first pass

\section{Related Work}

    Naturally, many solver developers author their own test suites to validate their solvers~\cite{cvc4-tests,z3str3-tests,z3str2-tests}. In addition, several popular problem suites are publicly available for solver validation, such as the Kaluza~\cite{kaluza} and Kausler~\cite{kausler} suites.

    There are likewise several fuzzers and problem generators currently available, but none of them can generate or transform string and regex problems. For example, the FuzzSMT~\cite{fuzzsmt} tool generates \smt{} problems with bit-vectors and arrays, but does not support strings or regexes. The SMTpp~\cite{smtpp} tool pre-processes and simplifies problems, but does not generate new ones or fuzz existing ones.

% FuzzSMT: http://fmv.jku.at/papers/BrummayerBiere-SMT09.pdf
% SMTpp: http://www.verit-solver.org/papers/smt2015.pdf