### 2018-01-30 includes/abstract.tex
cleaned up the paper a lot

\begin{abstract}
    In this paper, we introduce \fuzzer{}: a modular \smt{} problem
    transformer and generator for string solvers. We supply a
    repository of problems generated by \fuzzer{} in \smtfull{}
    format. We systemically compare \theSolvers{} on groups of such
    problems focusing on different combinations of string operators,
    and identify problems that are particularly challenging for some
    solvers. We briefly explain our observations and showcase how
    \fuzzer{} helped identify causes of performance degradations in
    \us{}.
\end{abstract}