### 2018-01-31 includes/problems.tex
First pass with Federico.

\section{Problem Suites}
\label{sec:problems}

In this section, we describe the benchmark suites that we generated
with \fuzzer{} on which the experimental evaluation was conducted.

\subsubsection{Generated Benchmarks}

We generated several suites
of string and regex problems using \fuzzer{}\footnote{For review
anonymity, we removed the link to the source code and the generated
suites. They will be provided separately in the artifact
submission.}. Table~\ref{tbl:generated} describes the suite
\texttt{generated} by \generator{}.
\begin{table}[H]
    \centering
    \caption{\generator{}-generated problem sets.}
    \label{tbl:generated}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}             & \textbf{Problems have ...} \\ \hline
        \textit{Concats}          & A single, deep tree of concats. \\ \hline
        \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
        \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
        \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
        \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
        \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
        \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
        \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
        \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
        \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
    \end{tabular}
\end{table}

\subsubsection{Transformed Benchmarks}

The \texttt{transformed}
suite contains benchmarks derived from existing seed problems by
iteratively applying \transformer{}. Table~\ref{tbl:transformed} shows
the overview. Every problem in the suite is named according to its
seed and the transformations it undertook. For example, input
\texttt{seed-fuzz-graft.smt2} was transformed by applying
\texttt{fuzz} and then \texttt{graft} from \texttt{seed.smt2}.
\begin{table}[H]
    \centering
    \caption{Transformed problem sets generated by \transformer{}.}
    \label{tbl:transformed}
    \begin{tabular}{|l|l|r|}
        \hline
        \textbf{Name}   & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
        \textit{Amazon} & Two industrial regex membership problems & 516\\ \hline
        \textit{Regex}  & Z3str2 regular expression test suite     & 10,577\\ \hline
    \end{tabular}
\end{table}

The Amazon category is obtained from two seeds: (1)
\texttt{z3-regex-1.smt2} encodes the assertions \texttt{(S not in
  aaabbb*)} and \texttt{(S in aaa*bbb*)} for a string constant S; (2)
\texttt{z3-regex-2-negative.smt2} includes regex \texttt{(S in
  aaabbb*)} and \texttt{(S not in aaa*bbb*)}.  Both problems come from
our industrial collaborators. In total, 516 unique problems were
generated from these two seeds. The Regex category is seeded by the
z3str2 regular expression test suite \cite{z3str2-tests}.  This test
suite contains 42 problems dealing with a wide variety of regular
expressions.  Through cumulative transformations, we have $10,577$
unique benchmarks.