### 2018-01-31 includes/experiment.tex
Removing commented-out sections. Fixing up more content.

\section{Experimental Results and Analysis}
\label{sec:data}

We generated several problem suites with \fuzzer{} that made one
solver perform poorly, but not others. These suites are
\theSuites{}. Figure~\ref{fig:cvc-hard} shows the suites that were
uniquely difficult for \cvc{}. Figure~\ref{fig:z3str3-hard} shows the
suites that were uniquely difficult for \us{}. All experiments were
run in series, with a timeout of 15 seconds, on the same computer running
Ubuntu Linux 16.04. The computer had 32GB of RAM and an
Intel\textregistered{} Core\texttrademark{} i7-6700 CPU with clock speed
of 3.40GHz.

\begin{figure}[h]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-extracts-small.png}
        \caption{Performance on concats-extracts-small}
        \label{fig:concats-extracts-small}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/different-prefix.png}
        \caption{Performance on different-prefix}
        \label{fig:different-prefix}
    \end{subfigure}
    \caption{Problems hard for \cvc{}}
    \label{fig:cvc-hard}

    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-balanced.png}
        \label{fig:concats-balanced}
        \caption{Performance on concats-balanced}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{data/graphs/concats-small.png}
        \label{fig:concats-small}
        \caption{Performance on concats-small}
    \end{subfigure}
    \caption{Problems hard for \us{}}
    \label{fig:z3str3-hard}
\end{figure}

\subsubsection{Usefulness to \us{}: A Case Study}

We found a number of performance-related issues and opportunities for
new heuristics in \us{} thanks to \fuzzer{}. For example, the
instances in the \textit{concats-big} suite generated by \fuzzer{}
helped us discover a missing heuristic. In particular, \us{} didn't
make full use of the solving context (e.g. some terms are empty
strings) to simplify the concatenations of a long list of string terms
before trying to reason about the equivalences among subterms. \us{}
therefore introduced a large number of unnecessary intermediate
variables and propagations. As an immediate consequence of
testing \us{} on \textit{concats-big} benchmarks, we were able to
identify the factors common to these instances and rapidly formulate a
hypothesis as to why they were performing poorly.