### 2018-01-31 includes/problems.tex

\section{Problem Suites}
\label{sec:problems}

In this section, we describe the benchmark suites that we generated
with \fuzzer{}, and on which we conducted our experimental
evaluation.\footnote{The link to the repository will be added after
double-blind review.}

Table~\ref{tbl:generated} describes the string and regex instances that were
generated by \generator{}. Table~\ref{tbl:transformed} lists benchmarks
derived from existing seed instances by iteratively applying \transformer{}.
Every transformed instance is named according to its
seed and the transformations it undertook. For example, input
\texttt{z3-regex-1-fuzz-graft.smt2} was transformed by applying
\textit{fuzz} and then \textit{graft} to \texttt{z3-regex-1.smt2}.

The \textit{Amazon} category contains 472 instances, derived from two seeds
supplied by our industrial collaborators. The \textit{Regex} category is
seeded by the z3str2 regular expression test suite~\cite{z3str2-tests}. This
test suite contains 42 instances dealing with a wide variety of regular
expressions. Through cumulative transformations, we expanded it to 7,551
instances. Finally, the \textit{Sanitizer} category is obtained from
five industrial e-mail address and IPv4 sanitizers.

\begin{table}[t]
    \centering
    \caption{\generator{}-generated problem sets.}
    \label{tbl:generated}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}             & \textbf{Instances have ...} \\ \hline
        \textit{Concats}          & A single, deep tree of concats. \\ \hline
        \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
        \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
        \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
        \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
        \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
        \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
        \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
        \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
        \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
    \end{tabular}
\end{table}

\begin{table}[t]
    \centering
    \caption{Transformed problem sets generated by \transformer{}.}
    \label{tbl:transformed}
    \begin{tabular}{|l|l|r|}
        \hline
        \textbf{Name}      & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
        \textit{Amazon}    & Two industrial regex membership problems & 472 \\ \hline
        \textit{Regex}     & Z3str2 regular expression test suite     & 7,551 \\ \hline
        \textit{Sanitizer} & 5 e-mail and IPv4 sanitiser examples     & 1,170 \\ \hline
    \end{tabular}
\end{table}