### 2018-01-31 includes/abstract.tex
Increasing consistency. Renaming problems.tex to suites.tex. Using the word 'instance' instead of the word 'problem'.

\begin{abstract}
    In this paper, we introduce \fuzzer{}: a modular \smt{} problem instance
    transformer and generator for string solvers. We supply a
    repository of instances generated by \fuzzer{} in \smtfull{}
    format. We systematically compare \theSolvers{} on groups of such
    instances, and identify instances that are particularly challenging for
    some solvers. We briefly explain our observations and show how
    \fuzzer{} helped identify causes of performance degradations in
    \us{}.
\end{abstract}