### 2018-01-31 includes/prior.tex

\section{Related Work}

Many solver developers author their own test suites to
validate their solvers~\cite{cvc4-tests,z3str3-tests,z3str2-tests}. In
addition, several popular instance suites are publicly available for
solver testing and benchmarking, such as the Kaluza~\cite{kaluza} and
Kausler~\cite{kausler} suites. There are likewise several fuzzers and
instance generators currently available, but none of them can generate
or transform string and regex instances. For example, the
FuzzSMT~\cite{fuzzsmt} tool generates \smt{} instances with bit-vectors
and arrays, but does not support strings or regexes. The
SMTpp~\cite{smtpp} tool pre-processes and simplifies instances, but
does not generate new ones or fuzz existing ones.