### 2018-01-31 includes/suites.tex
Adding more data to tables. Formatting tables for easier editing. Adding two more test problems.

\section{Instance Suites}
\label{sec:suites}

In this section, we describe the benchmark suites that we generated
with \fuzzer{}, and on which we conducted our experimental
evaluation.\footnote{The link to the repository will be added after
double-blind review.}

Table~\ref{tbl:generated} describes the string and regex instances that were
generated by \generator{}. Table~\ref{tbl:transformed} lists benchmarks
derived from existing seed instances by iteratively applying \transformer{}.
Every transformed instance is named according to its
seed and the transformations it undertook. For example, input
\texttt{z3-regex-1-fuzz-graft.smt2} was transformed by applying
\textit{Fuzz} and then \textit{Graft} to \texttt{z3-regex-1.smt2}.

The \textit{Amazon} category contains 472 instances, derived from two seeds
supplied by our industrial collaborators. The \textit{Regex} category is
seeded by the z3str2 regular expression test suite~\cite{z3str2-tests}. This
test suite contains 42 instances dealing with a wide variety of regular
expressions. Through cumulative transformations, we expanded it to 7,551
instances. Finally, the \textit{Sanitizer} category is obtained from
five industrial e-mail address and IPv4 sanitizers.

\begin{table}[t]
    \centering
    \caption{\generator{}-generated instances.}
    \label{tbl:generated}
    \begin{tabular}{|l|l|l|}
        \hline
        \textbf{Name}
            & \textbf{Instances have ...}
            & \textbf{Quantity}
            \\ \hline
        \textit{Concats}
            & A right-heavy, deep tree of concats.
            & 600
            \\ \hline
        \textit{Concats-Balanced}
            & A balanced, deep tree of concats.
            & 100
            \\ \hline
        \textit{Concats-Extracts}
            & A single concat tree, and random character extractions.
            & 600
            \\ \hline
        \textit{Lengths}
            & A single, large length constraint on a variable.
            & 1,000
            \\ \hline
        \textit{Lengths-Concats}
            & Variables and their concats, each with fixed lengths.
            & 500
            \\ \hline
        \textit{Overlaps}
            & A formula of the form A.X = X.B.
            & 100
            \\ \hline
        \textit{Regex}
            & A test for regex membership of arbitrary complexity.
            & 1,200
            \\ \hline
        \textit{Many-Regexes}
            & A test for membership in many random regexes.
            & 600
            \\ \hline
        \textit{Regex-Deep}
            & A test for membership in a regex with many nested operators.
            & 450
            \\ \hline
        \textit{Regex-Pair}
            & A test for membership in one regex, but not another.
            & 600
            \\ \hline
        \textit{Regex-Lengths}
            & A test for membership in a regex, and a length constraint.
            & 600
            \\ \hline
        \textit{Different-Prefix}
            & An equality between two deep concats with different prefixes.
            & 300
            \\ \hline
    \end{tabular}
\end{table}

\begin{table}[t]
    \centering
    \caption{Transformed instances generated by \transformer{}.}
    \label{tbl:transformed}
    \begin{tabular}{|l|l|r|}
        \hline
        \textbf{Name}      & \textbf{Seed}                             & \textbf{Quantity} \\ \hline
        \textit{Amazon}    & Two industrial regex membership instances & 472 \\ \hline
        \textit{Regex}     & Z3str2 regular expression test suite      & 7,551 \\ \hline
        \textit{Sanitizer} & 5 e-mail and IPv4 sanitiser examples      & 1,170 \\ \hline
    \end{tabular}
\end{table}