### 2018-01-31 includes/abstract.tex
Reverting some changes addressed in review. Made tables look scientific. Made sure tables were no longer than line width. Added footnotes rather than column for Ifaz transformations.

\begin{abstract}
    In this paper, we introduce \fuzzer{}: a modular \smt{} problem instance
    transformer and generator for string solvers. We supply a
    repository of instances generated by \fuzzer{} in \smtfull{}
    format. We systematically compare \theSolvers{} on groups of such
    instances, and identify those that are particularly challenging for
    some solvers. We briefly explain our observations and show how
    \fuzzer{} helped discover causes of performance degradations in
    \us{}.
\end{abstract}