### 2018-04-11 includes/suites.tex

\section{Instance Suites}
\label{sec:suites}

In this section, we describe the benchmark suites we generated
with \fuzzer{}, and on which we conducted our experimental
evaluation. Table~\ref{tbl:generated} describes the string
and regex instances that were
generated by \generator{}. Table~\ref{tbl:transformed} lists benchmarks
derived from existing seed instances by iteratively applying \transformer{}.
Every transformed instance is named according to its
seed and the transformations it undertook. For example,
\texttt{z3-regex-1-fuzz-graft.smt2} was transformed by applying
\textit{Fuzz} and then \textit{Graft} to \texttt{z3-regex-1.smt2}.

The \textit{Amazon} category contains 472 instances derived from two seeds
supplied by our industrial collaborators. The \textit{Regex} category is
seeded by the \usOld{} regular expression test suite~\cite{z3str2-tests}. This
test suite contains 42 instances dealing with a wide variety of regular
expressions. Through cumulative transformations we expanded the 42 seeds to 7,551 unique
instances. Finally, the \textit{Sanitizer} category is obtained from
five industrial e-mail address and IPv4 sanitizers.

\begin{table}[t]
    \caption{Repository of 10,258 \smtfull{} instances.}
    \begin{subtable}{1\textwidth}
        \centering
        \caption{\generator{}-generated instances.}
        \label{tbl:generated}
        \begin{tabular}{llr}
            \toprule
            \textbf{Name}
                & \textbf{Instances have a ...}
                & \textbf{Quantity}
            \\
            \midrule
            \textit{Concats-\{Small,Big\}}
                & Right-heavy, deep tree of concats.
                & 120 \\
            \textit{Concats-Balanced}
                & Balanced, deep tree of concats.
                & 100 \\
            \textit{Concats-Extracts-\{Small,Big\}}
                & Single concat tree, with character extractions.
                & 120 \\
            \textit{Lengths-\{Long,Short\}}
                & Single, large length constraint on a variable.
                & 200 \\
            \textit{Lengths-Concats}
                & Tree of fixed-length concats of variables.
                & 100 \\
            \textit{Overlaps-\{Small,Big\}}
                & Formula of the form A.X = X.B.
                & 80 \\
            \textit{Regex-\{Small,Big\}}
                & Complex regex membership test.
                & 120 \\
            \textit{Many-Regexes}
                & Multiple random regex membership tests.
                & 40 \\
            \textit{Regex-Deep}
                & Regex membership test with many nested operators.
                & 45 \\
            \textit{Regex-Pair}
                & Test for membership in one regex, but not another.
                & 40 \\
            \textit{Regex-Lengths}
                & Regex membership test, and a length constraint.
                & 40 \\
            \textit{Different-Prefix}
                & Equality of two deep concats with different prefixes.
                & 60 \\
            \bottomrule
        \end{tabular}
    \end{subtable}

    \begin{subtable}{1\textwidth}
        \centering
        \caption{\transformer{}-generated instances.}
        \label{tbl:transformed}
        \begin{tabular}{llr}
            \toprule
            \textbf{Name}      & \textbf{Seed}                              & \textbf{Quantity} \\
            \midrule
            \textit{Amazon}    & Two industrial regex membership instances. & 472 \\
            \textit{Regex}     & Z3str2 regular expression test suite.      & 7,551 \\
            \textit{Sanitizer} & Five e-mail and IPv4 sanitiser examples.   & 1,170 \\
            \bottomrule
        \end{tabular}
    \end{subtable}
\end{table}